Metadata collation

### 2023-09-22 metadata/parametadata_control.Rmd
Adding other files to metadata sheet

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggmap)
```

@@

```{r download raw data, eval=!file.exists("raw_data/mesoscope.gof")}
sum_url <- "http://scope.soest.hawaii.edu/data/scope2021/paragon1.sum"
download.file(sum_url, destfile = "raw_data/paragon1.sum")

gof_url <- "http://scope.soest.hawaii.edu/FTP/scope/water/paragon1.gof"
download.file(gof_url, destfile = "raw_data/paragon1.gof")

seaflow_url <- "http://scope.soest.hawaii.edu/data/scope2021/SeaFlow_dataset_v1.4.xlsx"
download.file(seaflow_url, destfile = "raw_data/SeaFlow_dataset_v1.4.xlsx", mode = "wb")

# register_google("AIzaSyBVNJV0Myoy3M50IBBDIjXZhUyU8Mnagvs")
# gm <- get_map(location = c(-156.5, 21.8), zoom = 8, maptype = "hybrid")
# saveRDS(object = gm, file = "raw_data/ggmap_of_aloha.rds")

magick::image_read("http://scope.soest.hawaii.edu/data/scope2021/Paragon1CTDchlp.pdf", density = 300) %>%
  magick::image_trim() %>%
  magick::image_write("raw_data/chl_pdf_url.png", format = "png")
```

![](raw_data/chl_pdf_url.png)

```{r From SCOPE FTP}
para_cast_times <- "raw_data/paragon1.sum" %>%
  read_table(skip = 3, col_names = FALSE, col_types = cols(.default = col_character())) %>%
  mutate(mdyhms=paste(X3, X4, X5, X6)) %>%
  mutate(time=as.POSIXct(strptime(mdyhms, "%b %d  %Y %X", tz="UTC"))) %>%
  select(station=X1, cast=X2, time) %>%
  mutate(station=as.numeric(station), cast=as.numeric(cast))

@@

# From ftp://ftp.soest.hawaii.edu/hot/water/Readme.water.jgofs
# Quality Indicators:
# Flag  Meaning
# 1    not quality controled
# 2    good data
# 3    suspect (i.e.  questionable) data
# 4    bad data
# 5    missing data
# 9    variable not measured during this cast
raw_gof <- read_table("raw_data/paragon1.gof", skip = 4, col_names = gof_names,
                      col_types = cols(.default = col_double()))
qual_df <- select(raw_gof, matches("^q\\d+$")) %>%
  unite("value", sep = "") %>%
  separate(sep = "", col = "value", into = c(NA, qual_names)) %>%
  mutate(across(everything(), as.numeric))
raw_gof[,qual_names][qual_df!=2] <- NA
para_core <- raw_gof %>% 
  left_join(para_cast_times) %>%
  select(station:rosette, time, lat:oxy_um, chl_ug:phaeo_ug) %>%
  mutate(lon=-lon)
```

@@

gm <- readRDS("raw_data/ggmap_of_aloha.rds")
stations_toplot <- para_core %>%
  distinct(station, cast, lat, lon) %>%
  mutate(station_cast=paste0("S", station, "C", cast)) %>%
  mutate(plotcol=ifelse(cast%in%my_casts, "red", "black"))
satmap <- ggmap(gm) +
  geom_path(aes(x=lon, y=lat), data=stations_toplot, color="grey80") +
  geom_point(aes(x=lon, y=lat), data = stations_toplot, color="white") +
  scale_color_identity()
statmap <- ggplot(stations_toplot, aes(x=lon, y=lat)) +
  geom_path(color="black") +
  geom_text(aes(x=lon, y=lat, color=plotcol, label=station_cast)) +
  scale_color_identity() +
  theme_bw() +
  coord_fixed()
cowplot::plot_grid(satmap, statmap, nrow=1)

@@

seaflowdata <- readxl::read_excel("raw_data/SeaFlow_dataset_v1.4.xlsx")
seaflowdata %>%
  mutate(time=str_remove(time, "\\+.*")) %>%
  mutate(time=strptime(time, format = "%Y-%m-%dT%H:%M:%S")) %>%
  mutate(time=as.POSIXct(time)) %>%
  select(time, starts_with("biomass")) %>%
  pivot_longer(-time) %>%
  filter(!(name=="biomass_synecho" & value>0.2)) %>%
  ggplot() +
  geom_point(aes(x=time, y=value)) +
  facet_wrap(~name, ncol=1, scales="free_y") +
  scale_x_datetime(timezone = "HST", date_breaks = "1 day", date_labels="%b %d")
```

@@

```{r}
samp_filedata <- "mzMLs" %>%
  list.files(pattern = "Smp.*mzML", recursive = TRUE, full.names = TRUE) %>% 
  data.frame(filepath=.) %>%
  mutate(filename=basename(filepath)) %>%
  mutate(shortname=str_remove(filename, "-wIS|-noIS-neg|-noIS-pos")) %>%
  mutate(shortname=str_remove(shortname, ".*_Smp_")) %>%
  mutate(IS_type=str_extract(filepath, "wIS|noIS")) %>%
  mutate(polarity=str_extract(filepath, "neg|pos")) %>%
  mutate(amendment=str_extract(shortname, "Nit|Amm|Arg|GMP")) %>%
  mutate(depth=str_extract(shortname, "Deep|Surf")) %>%
  mutate(startime=str_extract(shortname, "Eve|Morn")) %>%
  mutate(timepoint=str_extract(shortname, "T\\d+")) %>%
  mutate(tripl=str_extract(shortname, "[A-C](?=\\.mzML$)"))

@@

samp_filedata <- samp_filedata %>%
  left_join(station_samp_map)
```

```{r write out}
write_csv(samp_filedata, file = "metadata/samp_filedata.csv")
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggmap)
library(RaMS)
```

@@

```{r download raw data, eval=!file.exists("metadata/raw_data/paragon1.gof")}
sum_url <- "http://scope.soest.hawaii.edu/data/scope2021/paragon1.sum"
download.file(sum_url, destfile = "metadata/raw_data/paragon1.sum")

gof_url <- "http://scope.soest.hawaii.edu/FTP/scope/water/paragon1.gof"
download.file(gof_url, destfile = "metadata/raw_data/paragon1.gof")

seaflow_url <- "http://scope.soest.hawaii.edu/data/scope2021/SeaFlow_dataset_v1.4.xlsx"
download.file(seaflow_url, destfile = "metadata/raw_data/SeaFlow_dataset_v1.4.xlsx", mode = "wb")

# register_google("AIzaSyBVNJV0Myoy3M50IBBDIjXZhUyU8Mnagvs")
# gm <- get_map(location = c(-156.5, 21.8), zoom = 8, maptype = "hybrid")
# saveRDS(object = gm, file = "metadata/raw_data/ggmap_of_aloha.rds")

magick::image_read("http://scope.soest.hawaii.edu/data/scope2021/Paragon1CTDchlp.pdf", density = 300) %>%
  magick::image_trim() %>%
  magick::image_write("metadata/raw_data/chl_pdf_url.png", format = "png")
```

![](metadata/raw_data/chl_pdf_url.png)

```{r From SCOPE FTP}
para_cast_times <- "metadata/raw_data/paragon1.sum" %>%
  read_table(skip = 3, col_names = FALSE, col_types = cols(.default = col_character())) %>%
  mutate(mdyhms=paste(X3, X4, X5, X6)) %>%
  mutate(time=as.POSIXct(strptime(mdyhms, "%b %d  %Y %X", tz="UTC"))) %>%
  select(station=X1, cast=X2, time) %>%
  mutate(station=as.numeric(station), cast=as.numeric(cast))

@@

# From ftp://ftp.soest.hawaii.edu/hot/water/Readme.water.jgofs
# Quality Indicators:
# Flag  Meaning
# 1    not quality controled
# 2    good data
# 3    suspect (i.e.  questionable) data
# 4    bad data
# 5    missing data
# 9    variable not measured during this cast
raw_gof <- read_table("metadata/raw_data/paragon1.gof", skip = 4, col_names = gof_names,
                      col_types = cols(.default = col_double()))
qual_df <- select(raw_gof, matches("^q\\d+$")) %>%
  unite("value", sep = "") %>%
  separate(sep = "", col = "value", into = c(NA, qual_names)) %>%
  mutate(across(everything(), as.numeric))
raw_gof[,qual_names][qual_df!=2] <- NA
para_core <- raw_gof %>% 
  left_join(para_cast_times) %>%
  select(station:rosette, time, lat:oxy_um, chl_ug:phaeo_ug) %>%
  mutate(lon=-lon)
```

@@

gm <- readRDS("metadata/raw_data/ggmap_of_aloha.rds")
stations_toplot <- para_core %>%
  distinct(station, cast, lat, lon) %>%
  mutate(station_cast=paste0("S", station, "C", cast)) %>%
  mutate(plotcol=ifelse(cast%in%my_casts, "red", "black"))
satmap <- ggmap(gm) +
  geom_path(aes(x=lon, y=lat), data=stations_toplot, color="grey80") +
  geom_point(aes(x=lon, y=lat), data = stations_toplot, color="white") +
  scale_color_identity()
statmap <- ggplot(stations_toplot, aes(x=lon, y=lat)) +
  geom_path(color="black") +
  geom_text(aes(x=lon, y=lat, color=plotcol, label=station_cast)) +
  scale_color_identity() +
  theme_bw() +
  coord_fixed()
cowplot::plot_grid(satmap, statmap, nrow=1)

@@

seaflowdata <- readxl::read_excel("metadata/raw_data/SeaFlow_dataset_v1.4.xlsx")
seaflowdata %>%
  mutate(time=str_remove(time, "\\+.*")) %>%
  mutate(time=strptime(time, format = "%Y-%m-%dT%H:%M:%S")) %>%
  mutate(time=as.POSIXct(time)) %>%
  select(time, starts_with("biomass")) %>%
  pivot_longer(-time) %>%
  filter(!(name=="biomass_synecho" & value>0.2)) %>%
  ggplot() +
  geom_point(aes(x=time, y=value)) +
  facet_wrap(~name, ncol=1, scales="free_y") +
  scale_x_datetime(timezone = "HST", date_breaks = "1 day", date_labels="%b %d")
```

@@

```{r}
samp_filedata <- "mzMLs" %>%
  list.files(pattern = "Smp.*mzML", recursive = TRUE, full.names = TRUE) %>% 
  data.frame(filepath=.) %>%
  mutate(filename=basename(filepath)) %>%
  mutate(shortname=str_remove(filename, "-wIS|-noIS-neg|-noIS-pos")) %>%
  mutate(shortname=str_remove(shortname, ".*_Smp_")) %>%
  mutate(samp_type="Smp") %>%
  mutate(IS_type=str_extract(filepath, "wIS|noIS")) %>%
  mutate(polarity=str_extract(filepath, "neg|pos")) %>%
  mutate(amendment=str_extract(shortname, "Nit|Amm|Arg|GMP")) %>%
  mutate(depth=str_extract(shortname, "Deep|Surf")) %>%
  mutate(startime=str_extract(shortname, "Eve|Morn")) %>%
  mutate(timepoint=str_extract(shortname, "T\\d+")) %>%
  mutate(tripl=str_extract(shortname, "[A-C](?=\\.mzML$)"))

@@

other_filedata <- "mzMLs" %>%
  list.files(recursive = TRUE, full.names = TRUE) %>%
  str_subset("Smp.*mzML", negate = TRUE) %>%
  str_subset("MSMS", negate = TRUE) %>%
  str_subset("TMAO", negate = TRUE) %>%
  data.frame(filepath=.) %>%
  mutate(filename=basename(filepath)) %>%
  mutate(shortname=str_remove(filename, "-wIS|-noIS-neg|-noIS-pos")) %>%
  mutate(shortname=str_remove(shortname, ".*_Poo_|.*_Std_(4uMStds)?")) %>%
  mutate(samp_type=str_extract(filename, "Poo|Std")) %>%
  mutate(IS_type=str_extract(filepath, "wIS|noIS")) %>%
  mutate(polarity=str_extract(filepath, "neg|pos")) %>%
  mutate(amendment=str_extract(shortname, "Nit|Amm|Arg|GMP")) %>%
  fill(IS_type, amendment, .direction = "down")

all_filedata <- samp_filedata %>%
  bind_rows(other_filedata) %>%
  left_join(station_samp_map)
```

## Get file timestamps using RaMS

```{r}
mzml_timestamps <- grabMSdata(all_filedata$filepath, grab_what = "metadata") %>%
  pluck("metadata") %>%
  select(filename, timestamp)
parametadata <- left_join(all_filedata, mzml_timestamps)
```

## Write out

```{r write out}
write_csv(parametadata, file = "metadata/parametadata.csv")
```
